---
title: "Arctic-Data-Website-Demo"
author: "Priyanka Surio"
date: "August 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading Libraries

```{r, warning=FALSE, message=FALSE, echo=FALSE}
suppressPackageStartupMessages({
  library(leaflet)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(DT)
})
```

#Load Salmon Data from KNB

```{r, warning=FALSE, message=FALSE, echo=FALSE}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/knb.92020.1"
# data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"
esc <- tryCatch(
    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste(" 
                      Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        return(esc)
    }
)

head(esc)
```

#Calculating Medians

```{r, warning=FALSE, message=FALSE, echo=FALSE}
medianEsc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>%
  group_by(Species, SASAP.Region, Year, Location) %>%
  summarize(escapement = sum(DailyCount)) %>% 
  group_by(Species) %>% 
  summarise(medianEsc = median (escapement))


head(medianEsc)
```

#Plot escapement values as bar chart

```{r, warning=FALSE, message=FALSE, echo=FALSE}
ggplot(medianEsc, aes(Species, medianEsc)) +
  geom_col() +
  coord_flip()

```

#Building a map 

Starting from daily escapement data again and finding unique locations at which escapement was measured 


##Find unique location points

Deal with missing values in the location field

```{r, warning=FALSE, message=FALSE, echo=FALSE}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()
datatable(locations)
```


## Draw the map

Using leaflet to make an interactive map

```{r, warning=FALSE, message=FALSE, echo=FALSE}
leaflet(locations) %>%
  addTiles () %>% 
  addMarkers(~ Longitude, ~Latitude, popup = ~ Location)

```


Fixing the longitude values so they are not negative

```{r, warning=FALSE, message=FALSE, echo=FALSE}

locs <- locations %>%  mutate(Longitude = abs(Longitude)* -1)
leaflet(locs) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)
```

